---
title: "HW4"
author: "Mathieu Sauterey"
date: "October 9, 2018"
output:
  pdf_document: default
  word_document: default
---

# Problem 1

Stratified random sampling was used to estimate the total number of bushels of hard shell clams in Narragansett Bay, Rhode Island. The area of interest was divided into four strata based on preliminary surveys that identified areas in which clams were abundant. Then nh dredge tows were made in stratum h for h = 1, 2, 3, 4. The acreage for each stratum was known, and the area fished during a standard dredge tow was calculated to be 0.04 acres; thus we may use
$N_h$ = 25 x $Area_h$.

## (a) Estimate the total number of bushels of claims in the area, and give the standard error of your estimate. 

```{r}

area = c(222.81, 49.61, 50.25, 197.81)

n.h = c(4, 6, 3, 5)

N.h = round(25*area)

ybar.h = c(0.44, 1.17, 3.92, 1.80)

s.h = sqrt(c(0.068, 0.042, 2.146, 0.794))

t.hat_strat = sum(ybar.h*N.h)

V.hat.t <- sum(N.h^2 * s.h^2 / n.h * (1 - n.h/N.h))

SE.t <- sqrt(V.hat.t)

print(t.hat_strat)
print(SE.t)


```
E[$\hat{t}$] = 17,726.12

SE[$\hat{t}$] = 2,354.30

## (b) Another survey was performed at the end of the commercial season. In this survey, strata 1, 2, and 3 were collapsed into a single stratum, called stratum 1 below. Estimate the total number of bushels of clams (with standard error) at the end of the season.

```{r}

area = c(322.67, 197.81)

n.h = c(8, 5)

N.h = round(25*area)

ybar.h = c(0.63, 0.40)

s.h = sqrt(c(0.083, 0.046))

t.hat_strat = sum(ybar.h*N.h)

V.hat.t <- sum(N.h^2 * s.h^2 / n.h * (1 - n.h/N.h))

SE.t <- sqrt(V.hat.t)

print(t.hat_strat)
print(SE.t)

```
E[$\hat{t}$] = 7,060.21

SE[$\hat{t}$] = 948.28

# Problem 2

The data file agsrs contains information on the number of farms and acres devoted to farms, for an SRS of n = 300 counties from the population of N = 3078 in the United States. In 1987, the United States had a total of 2,087,759 farms.

## (a) Use ratio estimation to estimate the total number of acres devoted to farming in 1992, using the number of farms in 1987 as the auxiliary variable. Give a 95% confidence interval.

```{r}
# Loads the necessary package
library("SDaA")

# Loads the county data
county_data = agsrs

# Pop size
N = 3078
tx_U = 2087759

# Sample data and pop of x
x_87 = county_data$farms87
y_92 = county_data$acres92
xbar.U = tx_U / N

# Ratio estimator function 
ratio.estimator.mean <- function(x.samp, y.samp, N, xbar.U)
{
  n <- length(y.samp)
  xbar <- mean(x.samp); ybar <- mean(y.samp);
  B.hat <- ybar / xbar
  ybar.hat.r <- B.hat * xbar.U
  e <- y.samp - B.hat * x.samp
  V.hat <- (xbar.U/xbar)^2 * var(e)/n * (1 - n/N)
  SE <- sqrt(V.hat)
  answer <- c(point.est=ybar.hat.r, std.error=SE)
  return(answer)
}

# Calculates 1992 acres total + SE
result = ratio.estimator.mean(x.samp = x_87, y.samp = y_92, N, xbar.U)
N * result

# 95% CI for the population total of US acres in 1992
N * ( result[1] + c(-1,1) * 1.96 * result[2] )

```
E[$\hat{t}_{yr}$] = 960,155,061 acres

SE[$\hat{t}_{yr}$] = 68,446,406 acres

[ 826,000,106 ; 1,094,310,016 ] is a 95% CI for the total number of acres in the US devoted to farming in 1992.

## (b) Repeat part (a), using regression estimation.

```{r}

# Regression estimator function 
regression.estimator.mean <- function(x.samp, y.samp, N, xbar.U)
{
  n <- length(y.samp)
  xbar <- mean(x.samp); ybar <- mean(y.samp);
  fit <- lsfit(x.samp, y.samp)
  B1.hat <- as.numeric(fit$coefficients)[2]
  ybar.hat.reg <- ybar + B1.hat * (xbar.U - xbar)
  e <- fit$residuals
  V.hat <- var(e)/n * (1 - n/N)
  SE <- sqrt(V.hat)
  answer <- c(point.est=ybar.hat.reg, std.error=SE)
  return(answer)
}

# Calculates 1992 acres total + SE
result <- regression.estimator.mean(x.samp = x_87, y.samp=y_92, N=N, xbar.U=xbar.U)
N * result

# 95% CI for the population total of US acres in 1992
N * ( result[1] + c(-1,1) * 1.96 * result[2] )

```
E[$\hat{t}_{yreg}$] = 921,406,265 acres

SE[$\hat{t}_{yreg}$] = 58,065,813 acres

[ 807,597,271 ; 1,035,215,259 ] is a 95% CI for the total number of acres in the US devoted to farming in 1992.


# Problem 3

## Use the data in agsrs to estimate the total number of acres devoted to farming in 1992 for each of two domains:

## (a) counties with fewer than 600 farms.

```{r}

# Sets up domains
domain.samp <- ifelse(county_data$farms92 < 600, 1 , 0)

# Domain estimation function
domain.estimation <- function(y.samp, domain.samp, d, N)
{
 n <- length(y.samp)
 x <- (domain.samp==d); y <- y.samp; u <- x*y;
 u.bar <- mean(u); s2.u <- var(u);
 t.yd.hat <- N * u.bar
 V.hat <- N^2 * s2.u/n * (1 - n/N)
 SE <- sqrt(V.hat)
 answer <- c(point.est=t.yd.hat, std.error=SE)
 return(answer)
}

# Calculates domain total (1992 acres total + SE for domain farms < 600)
domain.estimation(y.samp = y_92, domain.samp, d = 1, N)

```
E[$\hat{t}_{yr}$] = 497,939,808 acres

SE[$\hat{t}_{yr}$] = 55,919,525 acres

(b) counties with 600 or more farms

```{r}

# Calculates domain total (1992 acres total + SE for domain farms >= 600)
domain.estimation(y.samp = y_92, domain.samp, d = 0, N)

```

E[$\hat{t}_{yr}$] = 418,987,302 acres

SE[$\hat{t}_{yr}$] = 38,938,277 acres
